### 2024-08-13 drafts/draftAug2024.qmd
started writing with quarto

---
title: "**Title**: Linking pinyon jay population abundance to pinyon cone production and climate in the SW United States"
format:
  docx:
    reference-doc: custom-reference-doc.docx
editor: visual
bibliography: references.bib
csl: journal-of-animal-ecology.csl
---

**Author list**: Ana Miller-ter Kuile^1,2^, Andreas Wion^3^, Jamie S. Sanderlin^2^

**Author affiliations**:

1 Northern Arizona University School of Informatics, Computing, and Cyber Systems, Flagstaff, Arizona, USA

2 USDA Forest Service Rocky Mountain Research Station, Flagstaff, Arizona, USA

3 USGS Santa Fe, New Mexico, USA

**Keywords**:

# Data availability

\newpage

# Introduction

Based on data from [@abadi2010]

# Methods

$$
ldsjfsdlkj
$$ {#eq-a}

\begin{equation}\\tag{1}\\label{eq-a}
a
\end{equation}

# Results

# Discussion

# References

```{r}
#| echo: false
#| eval: false
#| include: false

#found info on adding reference styles here:
#https://quarto.org/docs/authoring/citations.html
#and list of csls here:
#https://www.zotero.org/styles

```

# Tables and Figures