### 2025-04-29 meetings/april2025_updates.qmd
updating code for model to get oos data prepped and also RMSE for both in and out of sample

---
title: "Model update for: Linking pinyon jay population abundance to pinyon cone production and climate in the SW United States"
format: html
editor: visual
bibliography: references.bib
toc: true
embed-resources: true
---

```{r}
#| echo: false
#| include: false
#| warning: false
#| message: false

package.list <- c("here", "tidyverse")

## Installing them if they aren't already on the computer
new.packages <- package.list[!(package.list %in% installed.packages()[,"Package"])]
if(length(new.packages)) install.packages(new.packages)

## And loading them
for(i in package.list){library(i, character.only = T)}

source(here('code',
       '03_visualization',
       'weight_corrected_covarate_effects.R'))

```

## Project collaborators

Ana Miller-ter Kuile, Andreas Wion, Kyle Rodman, Kiona Ogle, Jamie Sanderlin

## Project goal

This project links data on modeled cone availability in the Southwest USA [@wion2025] to abundance of pinyon jays from eBIRD. Our aim is to explore the relationship between cone and bird abundance, including interactions with other mediating variables and also look at the temporal signal of cone-bird relationships.

### Covariates in the model

The covariates we're considering include:

-   cone abundance (with yearly lead and lag effects)

-   maximum temperature (with seasonal lag effects)

-   precipitation (with seasonal lag effects)

-   monsoonality (percent of 30-year average precipitation occurring July & August)

-   pinyon basal area (yearly)

We also considered the interactions with all covariates and cone availability, since we were interested in how these variables influenced the relationship with cone availability.

```{r}
#| echo: false
#| message: false
#| warning: false
#| fig-cap: "Figure 1: Temporal hypotheses for cone-jay abundance relationship. Birds could respond in a combination of time periods, including before or after mature cones are available."

knitr::include_graphics(here('pictures',
                             'time_conceptual.jpg'))
```

### Structure of the model

This model incorporates observed data from eBIRD and accounts for detection error in observations. Thus, we are modeling true latent abundance, *N*, based on partially observed data from eBIRD with detection error and survey covariates.

We spatially and temporally stratified checklists in our analyses based on an adaption of the [best practices](https://cornelllabofornithology.github.io/ebird-best-practices/abundance.html) outlined by Cornell Labs for using eBIRD data for science. We spatially sub-sampled based on a 4x4 km grid (since this was the scale of all our covariates). We incorporated spatial uncertainty by then buffering each checklist within that grid by 1/2 of its sampling distance from eBIRD and then merging all checklists in a 4x4 km grid into a "blob" that represented the spatial unit for estimating jay relationships with covariates and expected abundance, $\lambda$ (checklists = repeat observations within a site/year). Within the blob, we determined the amount of the blob overlapping the original 4x4 km grid cell and all surrounding cells and used covariates calculated as the weighted average of covariate values in each cell times the amount of overlap of the blob with that cell. For every checklist in a blob, we determined "true" abundance, *N*, by multiplying the blob-level rate parameter, $\lambda$, in the Poisson distribution by the area surveyed by each checklist in that blob. Thus, *N* is estimated at the checklist level and shares biological information across checklists within a blob.

You can examine the model code [here](https://github.com/anamtk/pinyon_jays/blob/main/code/jags_models/ebird_abund_JAGS_checkN_KO.R).

## Results

### Main effects of covariates

Here are the overall results for covariate effects (I know they're missing 95% BCI - have to calculate after the fact because reasons, and haven't gotten to it yet! All are non-zero except maybe ConexBasal Area interaction). All are scaled and so their magnitudes can be compared.

```{r}
#| echo: false
#| message: false
#| warning: false
#| fig-cap: "Figure 2: Covariate effects for the pinyon jay abundance model with eBIRD data. These include overall effects of covariates as well as their interactions with cone abundance."

covariate_betas

```

In a nutshell:

-   More birds with more cones (yearly lags)

-   A stronger positive effect of cones with more precipitation (seasonal lags)

-   A stronger positive effect of cones with higher maximum temperatures (seasonal lags)

-   More birds in areas with more monsoon moisture

-   More birds at higher pinyon basal area

-   (Weak) Stronger positive relationship with cones when pinyon basal area is higher

-   A weaker positive relationship with cones as monsoonality increases

-   Fewer birds at higher maximum temperatures (seasonal lags)

-   Fewer birds at higher precipitation (seasonal lags)

### Cone temporal effects

```{r}
#| echo: false
#| message: false
#| warning: false
#| fig-cap: "Figure 3: Temporal effect of cone abundance effect on pinyon jays. Jays respond in abundance the spring before summer/fall cone initiation (2 years before cones mature) and also in the year following cone maturation and three years following cone maturation."

cone_weight_plot

```

In a nutshell:

-   Birds increase in abundance the spring before cones initiate in the fall/summer. Something about climate triggers also seem to help birds?

-   Birds increase the year following cones, suggesting a caching effect

-   Birds increase three years following cones - could this be some kind of delayed maturation effect? So - fledglings from the big cache year mature two years later and enter breeding population?

### Climate temporal effects

```{r}
#| echo: false
#| message: false
#| warning: false
#| fig-cap: "Figure 4: Temporal effect of maximum temperature and precipitation (both have negative effects on jay abundance). Seasons are based on bird biology: Br = breeding (February-April); Fl = adults with dependent young (May-June), Su = summer (unknown behavior; July); SW = summer/winter dispersed foraging (this is when they travel all over the place; August-January)"

clim_weights_plot

```

In a nutshell:

-   It does seem like precipitation is most important during breeding and raising dependent young (generally). All three previous periods with dependent young are important (wetter during those periods = fewer birds); two and three year ago breeding seasons also important; second mystery time summer also important. **In all these periods, when they are wetter, there are fewer birds.**

-   Temperature seems most important \~2 years ago. Two years ago winter is important, as is the breeding and dependent young period two years ago. (need to read about this!) **Again, these indicate that when these periods are warmer, there are fewer birds.**

-   Again, I'm wondering if this two to three year lag in birds is related to when recently-fledged individuals become breeding adults?

### Interactions

I haven't had a chance to spend too much time in the interactions, beyond what I've highlighted above about their main effects. However, these also include temporal effects, so they have sort of a "this, then this, then this" type relationship with cones and birds. For example, the positive interaction between cones and precipitation means that birds rely more on cones in wetter periods. Two of the most important wet periods are the breeding and fledging seasons (spring/early summer) two years ago. Thus, when there is a wet spring/early summer followed by a cone initiation year, there will be more birds in areas where cones are maturing both in that year and in the year following cone maturation (2 years later). Maybe it's more important to focus just on the main effects rather than get bogged down in the temporal aspects of the interactions... not sure. But given that wet breeding seasons 2 years ago lead to fewer birds and that birds rely more on cones in wet times, maybe the more important factor here is that cones can offset some of the negative effects, or that birds adapt their reliance on cones to be more heavily dependent on them when other conditions are less favorable. That jays depend less on cones in areas of the SW with more monsoonal moisture suggests that these places potentially have more other foods available. Conversely, these are areas (like in NM) where there's been a lot of piñon die-off, so that may be related too.